# Analysis on storm aganistpopulation healh



## 1- Synopsis
  - This paper presents an analysis about the effect of storm on the population health.[Why we study this topic]. We analysis the data published by the )National Weather Service Instruction in 2007. These data . We use [data process technic]. We found[research findins] . [Implication and advice]


## 2 - Analysis Question
 1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to popuulation health
 2.  Across the United States, which types of events have the greatest economic consequences
 
## 3- Data Processing

- 1. Load data
```` {r echo = TRUE,cache=TRUE}
#attributes
#dUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
#download.file(dUrl, f, mode = "wb")
ff <- file.path(getwd(), "repdata-data-StormData.csv.bz2")
data <- read.csv(ff, stringsAsFactors = FALSE,quote='')
````

- 2. Variables we keep for analysis
  Variabl Name | Description|
  -------------|-------------|
   state      | Abbervation for states in US |
   evtype     | Event type |
   fatalites  | The number of people died|
   injures   | The number of people injuured|
   propdmg    | The amount of property damge (measured in money)|
   propdmgexp  | The unit of  damge (B,M,K,H)|
   corpdmg  | The amount of corp damge (measured in money)|
   corpdmgexp    | The unit of  damge (B,M,K,H)|
  
```` {r filterCols,echo = TRUE}

#attributes<-c('X.STATE.','X.EVTYPE.','X.FATALITIES.','X.INJURIES.','X.PROPDMG.','X.PROPDMGEXP.','X.CROPDMG##.','X.CROPDMGEXP.')
attributes<-c('STATE','EVTYPE','FATALITIES','INJURIES','PROPDMG','PROPDMGEXP','CROPDMG','CROPDMGEXP')
pData<-data[,attributes]
colnames(pData)<-c('state','evtype','fatalites','injures','propdmg','propdmgexp','corpdmg','corpdmgexp')
````
### 3.1 Data cleanning
   
#### 3.1.1  Event tpye 
      - Create a column as the unified name of that eventype 
      - The hierachy looks like as follow:
        - Convection
          - lightning
          - tornado
          - thunderstorm Wind
          - hail
        - ExtremeTemp
          - cold
          - heat
        - Marine
          - costal Storm
          - sunami
          - rip Current
        - Flood
          - flash Flood
          - river Flood
        - Tropical Cyclones
          - tropical Storm
          - hurricane
        - Other
          - drought
          - dust Storm
          - dust Devil
          - rain
      - The unified event name and filtr expression:
 |  Event    |    Filter expression  |  
------------------| ----------------|
  Lightning |  LIGHTNING | 
  Tornado   | NADO,FUNNEL,WATERSPOUT"|
  Thunderstorm wind |  THUNDER,STORM,WIND | 
  Hail |  HAIL | 
  Cold |  BLIZZARD|WINTER|COLD|LOW TEMP|RECORD LOW|SNOW|ICE | 
  Heat |  HEAT|WARM|RECORD HIGH | 
  Costa l Storm |  COSTAL STORM | 
  Sunami | SUNAMI | 
  Rip current|RIP CURRENT
  Flash flood   |  FLASH FLOOD,FLD,FLOOD | 
  River flood   |  RIVER FLOOD , URBAN FLOOD | 
  Tropical Storm |TROPICAL STORM,TROPICAL|
  Hurricane   | HURRICANE |
   Drought   | DROUGHT |
    Dust storm   | DUST STORM |
        Dust devil   | DUST DEVIL |
    RAIN  | rain |


```` {r unifiednamefun,cache=TRUE,echo = TRUE}
deliverEvtUnifiedName<-function (dd){
  evtMatcher<-data.frame( reg=c("NADO|FUNNEL|WATERSPOUT","THUNDER|STORM|WIND","HAIL"
                                ,"FROST|FREEZ|BLIZZARD|WINTER|COLD|LOW TEMP|RECORD LOW|SNOW|ICE","HEAT|WARM|RECORD HIGH"
                                ,"COSTAL STORM","SUNAMI","RIP CURRENT"
                                ,"FLASH FLOOD|FLD|FLOOD","RIVER FLOOD|URBAN FLOOD","TROPICAL STORM|TROPICAL"," "
                                ,"DROUGHT","DUST STORM","DUST DEVIL","RAIN","LIGHTNING") )
  
  factorId<-c("Tornado","Thunderstorm wind","Hail"
                        ,"Cold","Heat","Costal Storm","Sunami","Rip current"
                        ,"Flash flood","River flood","Tropical Storm","Hurricane"
                        ,"Drought","Dust storm","Dust devil","Rain","Ligntning","Other");
 
  for ( i in  1:nrow(evtMatcher)){
    indexFit<-grep(evtMatcher[i,"reg"],toupper(dd[ ,"evtype"]))
    if(length(indexFit)>0){
      dd[indexFit,"event"]<- as.character(factorId[i])
     
    }
  }
}
```

  - Use the function to filter and put value into "event"
```` {r cache=TRUE,echo = TRUE}


pData$event<-("-")

# Give each event an unified id factor (17+1 factors)
pData<-deliverEvtUnifiedName(pData)
otherIndex<-grep("-",pData[,"event"])
pData[otherIndex,"event"]<-"Other"
pData$event<-as.factor(pData$event)

# See the event and number of that
table(sort(-pData$event))
````

#### 3.1.2  Create a variable has exact amount for property and corp damage
        - Fomula I  : propdmgValue  = propdmg       *  propdmgexp
        - Fomula II : corpdmgValue  = cropdmg       *  cropdmgexp
        - New Variable: ecodmgvalue = propdmgValue+corpdmgValue
        
              propdmgexp   |  unit   
           ----------------| ----------------
              B            |  1,000,000,000
              M            |      1,000,000   
              K            |          1,000 
              H            |            100
              NA or BLANK  |              1

```` {r dmgCaculation,cache=TRUE,echo = TRUE} 
deliverActualDmgValue<-function (dd){
  unit<-data.frame( cha=c("B","M","K","H"), val=c(1000000000,1000000,1000,100))
  dd$cropdmgvalue<-as.numeric(dd$corpdmg)
  dd$propdmgvalue<-as.numeric(dd$propdmg)
  for ( i in  1:nrow(unit)){
    indexCrodFit<-grep(as.character(unit[i,"cha"]),toupper(dd[ ,"corpdmgexp"]))
    if(length(indexCrodFit)>0){
      dd[indexCrodFit,'corpdmgvalue']<- unit[i,"val"]* dd[indexCrodFit ,"corpdmg"]
    }
    indexProdFit<-grep(as.character(unit[i,"cha"]),toupper(dd[ ,"propdmgexp"]))
    if(length(indexProdFit)>0){
      dd[indexProdFit,'propdmgvalue']<- unit[i,"val"]* dd[indexProdFit ,"propdmg"]
    }
  }
  return (dd)
}

pData<-deliverActualDmgValue(pData);

# see the value:
summary(pData$corpdmgvalue)
summary(pData$prodmgvalue)


# Add a new variable: propdmagevalue + corpdmgvalue
pData$ecodmgvalue<-pData$corpdmgvalue+pData$propdmgvalue
pData$pephealthdmg<-pData$injures+pData$fatalites
summary(pData$ecodmgvalue)
````
          
#### 3.2 Data Aggeratiion
```` {r dataAggrate,cache=TRUE,echo = TRUE} 

t<-aggregate(cbind(propdmgvalue, corpdmgvalue,injures,fatalites) ~ event, pData, FUN = sum)

tidy <- melt(t, id.var = "event", variable.name = "variable")
print(tidy)
````

### 4- Results
#### 4.1  The most harmful for people health top 5

````{r plot_healthdmg,cache=TRUE} 
damgeOnHealth<-tidy[ grep('fatalites|injures',tidy[,'damagetype'] ),]

ggplot(damgeOnHealth, aes(x=reorder(event,-value) ,y=value,fill=damagetype ))+geom_bar(stat='identity')+labs(title = "Top harmful weather event for populaiton health", x = "Event", y = "Damage on populaiton health (number of people")+ scale_fill_manual(values=c("#EED094", "#C70000"))
````
 Rank| Event   |  People Headth Harm (injuries+fatalites)| 
 -|--------|-------------|
1|           Tornado     |   97078 | 
 2|Thunderstorm wind    |    12608 | 
 3|             Heat    |    12400 | 
 4|       Flash flood   |     10250| 
 5|      Cold   |     7415| 



#### 4.2  The most harmful for economic value top 5
 Rank| Event   |  People Headth Harm (injuries+fatalites)| 
 -|--------|-------------|
 1|           Tornado     |   97078 | 
 2|Thunderstorm wind    |    12608 | 
 3|             Heat    |    12400 | 
 4|       Flash flood   |     10250| 
 5|      Cold   |     7415| 



```` {r plot_ecodmg,cache=TRUE} 

damgeOnEconomic<-tidy[ grep('corpdmgvalue|propdmgvalue',tidy[,'damagetype'] ),]

ggplot(damgeOnEconomic, aes(x=reorder(event,-value) ,y=value,fill=damagetype ))+geom_bar(stat='identity')+labs(title = "Top harmful weather event for economic ", x = "Event", y = "Damage on economic value (US dollors)")+ scale_fill_manual(values=c("#1E87DE", "#4D7A49"))

install.packages("xtable")
library(xtable)
sortedTable <- xtable(sort(tidy))
print(sortedTable, type = "html")
````
