# Analysis on Various Weather Events aganist Population Health and Economy

## 1- Synopsis
This report presents an analysis about the impacts of wheather events on the population health and economic. We analysed the data published by the National Weather Service Instruction in 2007. We will address the damge on economic  by reporting the economic dmange caculated by property damge and crop damge measuered in US dollars, and address the damage on people health by sum the number of injures and fatalites in the events. In the end, we illustrate the impact of different types of events on these two aspects with two barchart containing two measurements for each aspect. We found that the tornado is most harmful for population heath, and the flash flood results in greatest loss on economic value.

## 2 - Analysis Question
 1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to popuulation health
 2.  Across the United States, which types of events have the greatest economic consequences
 
## 3- Data Processing

- 1. Load data
```` {r echo = TRUE,cache=TRUE}

library(ggplot2) # plot 


````
```` {r echo = TRUE,cache=TRUE}
#attributes

ff <- file.path(getwd(), "repdata-data-StormData.csv.bz2")
data <- read.csv(ff, stringsAsFactors = FALSE)
````

- 2- Variables we keep for analysis

  Variabl Name | Description|
  -------------|-------------|
   state      | Abbervation for states in US |
   evtype     | Event type |
   fatalites  | The number of people died|
   injures   | The number of people injuured|
   propdmg    | The amount of property damge (measured in money)|
   propdmgexp  | The unit of  damge (B,M,K,H)|
   cropdmg  | The amount of corp damge (measured in money)|
   cropdmgexp    | The unit of  damge (B,M,K,H)|


```{r filterCols}
#attributes

attributes<-c('STATE','EVTYPE','FATALITIES','INJURIES','PROPDMG','PROPDMGEXP','CROPDMG','CROPDMGEXP')

pData<-data[,attributes]
colnames(pData)<-c('state','evtype','fatalites','injures','propdmg','propdmgexp','cropdmg','cropdmgexp')
names(pData)
```
### 3.1 Data cleanning
   
#### 3.1.1  Event tpye 

      - Create a column as the unified name of that eventype 
      - The hierachy looks like as follow: 
        - Convection
          - lightning
          - tornado
          - thunderstorm Wind
          - hail
        - ExtremeTemp
          - cold
          - heat
        - Marine
          - costal Storm
          - sunami
          - runami
          - rip Current
        - Flood
          - flash Flood
          - river Flood
        - Tropical Cyclones
          - tropical Storm
          - hurricane
        - Other
          - drought
          - dust Storm
          - dust Devil
          - rain
          
#### The unified event name and filtr expression:
  Event    |    Filter expression  |  
------------------| ----------------|
  Lightning |  LIGHTNING | 
  Tornado   | NADO,FUNNEL,WATERSPOUT"|
  Thunderstorm wind |  THUNDER,STORM,WIND | 
  Hail |  HAIL | 
  Cold |  BLIZZARD|WINTER|COLD|LOW TEMP|RECORD LOW|SNOW|ICE | 
  Heat |  HEAT|WARM|RECORD HIGH | 
  Costa l Storm |  COSTAL STORM | 
  Sunami | SUNAMI | 
  Rip current|RIP CURRENT
  Flash flood   |  FLASH FLOOD,FLD,FLOOD | 
  River flood   |  RIVER FLOOD , URBAN FLOOD | 
  Tropical Storm |TROPICAL STORM,TROPICAL|
  Hurricane   | HURRICANE |
  Drought   | DROUGHT |
  Dust storm   | DUST STORM |
  Dust devil   | DUST DEVIL |
  Rain  | RAIN |


```` {r unifiednamefun,cache=TRUE,echo = TRUE}
# Function : Give each eventype an unified event description
deliverEvtUnifiedName<-function (dd){
  evtMatcher<-data.frame( reg=c("NADO|FUNNEL|WATERSPOUT","THUNDER|STORM|WIND","HAIL"
                                ,"FROST|FREEZ|BLIZZARD|WINTER|COLD|LOW TEMP|RECORD LOW|SNOW|ICE",
                                "HEAT|WARM|RECORD HIGH","COSTAL STORM","SUNAMI","RIP CURRENT"
                                ,"FLASH FLOOD|FLD|FLOOD","RIVER FLOOD|URBAN FLOOD","TROPICAL STORM|TROPICAL"
                                ,"HURRICANE","DROUGHT","DUST STORM","DUST DEVIL","RAIN","LIGHTNING") )
  
  factorId<-c("Tornado","Thunderstorm wind","Hail"
                        ,"Cold","Heat","Costal Storm","Sunami","Rip current"
                        ,"Flash flood","River flood","Tropical Storm","Hurricane"
                        ,"Drought","Dust storm","Dust devil","Rain","Ligntning");
 
  for ( i in  1:nrow(evtMatcher)){
    indexFit<-grep(evtMatcher[i,"reg"],toupper(dd[ ,"evtype"]))
    if(length(indexFit)>0){
      dd[indexFit,"event"]<- factorId[i]
     
    }
  }
  return (dd)
}
```

#### Use the function to filter and put value into "event"
```` {r eventProcessing,echo = TRUE}


pData$event<-("-")

# Give each event an unified id factor (17+1 factors)
pData<-deliverEvtUnifiedName(pData)
otherIndex<-grep("-",pData[,"event"])
pData[otherIndex,"event"]<-"Other"
pData$event<-as.factor(pData$event)

# See the event and number of that
table(pData$event)
````


### 3.1.2  Create an exact amounts for property and corp damage

  - Fomula I  : propdmgValue = propdmg       *  propdmgexp
  - Fomula II : cropdmgValue = cropdmg       *  cropdmgexp  

        
              propdmgexp   |  unit   
           ----------------| ----------------
              B            |  1,000,000,000
              M            |      1,000,000   
              K            |          1,000 
              H            |            100
              NA or BLANK  |              1

#### Function : Caculate the property and crop damg by 'dmg' and 'exp' variable
  
```` {r dmgCaculation ,echo = TRUE} 
# Function : Caculate each entry the property and crop damge value by 'dmg' and 'exp' variables
deliverActualDmgValue<-function (dd){
  unit<-data.frame( cha=c("B","M","K","H"), val=c(1000000000,1000000,1000,100))
  multi<-c(1000000000,1000000,1000,100);

  for ( i in  1:nrow(unit)){
    # index that match the unit
    indexCrodFit<-grep(unit[i,"cha"], toupper(dd[ ,"cropdmgexp"]))
    
    if(length(indexCrodFit)>0){
      
      # Caculate the actual value
      dd[indexCrodFit,'cropdmgvalue']<- multi[i] * dd[indexCrodFit ,"cropdmg"]
      
    }
    
    # Same procudure for property damage
    indexProdFit<-grep(unit[i,"cha"],toupper(dd[ ,"propdmgexp"]))
    
    if(length(indexProdFit)>0){
      
      dd[indexProdFit,'propdmgvalue']<- multi[i] * dd[indexProdFit ,"propdmg"]
    }
    
  }
  return(dd)
}

# Default value of the damage equals to variable 'dmg'
pData$cropdmgvalue<-pData$cropdmg
pData$propdmgvalue<-pData$propdmg

pData<-deliverActualDmgValue(pData)
```
#### Watch the summary of crop damage and property damage
````{r summary1,echo = TRUE} 

# see the value:
summary(pData$cropdmgvalue)
summary(pData$propdmgvalue)

````

### Add  new variables: Damge on economic, Damge on population

 - Damge on economic : `ecodmgvalue` = `propdmagevalue` + `corpdmgvalue`
 - Damge on population health : `pephealthdmg` = `propdmagevalue` + `corpdmgvalue`
    
````{r addnewvariable,echo = TRUE} 
pData$ecodmgvalue<-pData$cropdmgvalue+pData$propdmgvalue
pData$pephealthdmg<-pData$injures+pData$fatalites
summary(pData$ecodmgvalue)
````
          
### 3.2 Data Aggeratiion

Caculate the Sum of `propdmgvalue`, `cropdmgvalue`,`injures`,`fatalites` for each `event` 

```` {r dataAggrate,echo = TRUE} 
library(reshape2)

t<-aggregate(cbind(propdmgvalue, cropdmgvalue,injures,fatalites) ~ event, pData, FUN = sum)
tidy <- melt(t, id.var = "event", variable.name = "variable")
colnames(tidy)<-c("event","damagetype","value")
print(tidy)
````

## 4- Results
### 4.1  The most harmful weather events for people health
  - Extract rows that contains info about people impact on populaiton health
````{r plot_healthdmg,fig.width=10} 
library("ggplot2")
damgeOnHealth<-tidy[ grep('fatalites|injures',tidy[,'damagetype'] ),]

ggplot(damgeOnHealth, aes(x=reorder(event,value) ,y=value,fill=factor(damagetype,labels=c("Injures","Fatalites") )))+geom_bar(stat='identity')+labs(title = "Top harmful weather event for populaiton health", x = "Event", y = "Damage on populaiton health (number of people)")+ scale_fill_manual(values=c("#F7B388", "#C70000"))+guides(fill = guide_legend(title = "Type of damage"))+xlab("Weather Event")+theme(axis.text = element_text(size=12,colour = "#1F4178"),                                                                                                                                                                                                                                                                                                                                                                                                              axis.title=element_text(size=14,colour = "#3A3E42",face="bold"),title = element_text(size=16,colour = "#282B2E",face="bold"))+coord_flip()
````

 
### 4.2  The most harmful weather events for economic value 

 - Extract rows that contains info about people impact on economy 
 
```` {r plot_ecodmg,fig.width=10} 
damgeOnEconomic<-tidy[ grep('cropdmgvalue|propdmgvalue',tidy[,'damagetype'] ),]
ggplot(damgeOnEconomic, aes(x=reorder(event,value) ,y=value,
                            fill=factor(damagetype,labels=c("Property","Crop") )))+geom_bar(stat='identity')+labs(title = "Top harmful weather event for economy ", x = "Event", y = "Damage on economic value (US dollors)")+ scale_fill_manual(values=c("#7D5E35", "#407D47"))+ guides(fill = guide_legend(title = "Type of damage"))+xlab("Weather Event")+theme(axis.text = element_text(size=12,colour = "#1F4178"),                                                                                                                                                                                                                                                                                                                                                                                                              axis.title=element_text(size=14,colour = "#3A3E42",face="bold"), title = element_text(size=16,colour = "#282B2E",face="bold")  )+ coord_flip()                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                        
````

## 5- Conclusion
  From the results, we can see :
  - 1. The most harmful weather event for population health is ***tornado*** 
  - 2. The most harmful weather event for economy is ***flash flood***




